### 2022-02-01 Writing/manuscript_gwr.rmd
writing folder created

---
title: "Economic Evaluation of Geographically Weighted Regression Analysis for Site-specific Nitrogen Management"
author:
  - Taro Mieno^[University of Nebraska Lincoln, tmieno2@unl.edu], Xiaofei Li^[Mississippi State University, xiaofei.li@msstate.edu], David S. Bullock^[University of Illinois, dsbulloc@illinois.edu]
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: true
    reference_docx: "word_template.docx"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ": "
bibliography: PA.bib
csl: field-crops-research.csl
abstract: ""
---

```{r echo = F, cache = F}
suppressMessages(library(knitr))
suppressMessages(library(here))
suppressMessages(library(officedown))
suppressMessages(library(officer))

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = F,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(lfe)
library(modelsummary)
library(patchwork)
```

```{r figure_setup, cache = F}
theme_update(
  axis.title.x =
    element_text(
      size = 12, angle = 0, hjust = .5, vjust = -0.3, face = "plain"
    ),
  axis.title.y =
    element_text(
      size = 12, angle = 90, hjust = .5, vjust = .9, face = "plain"
    ),
  axis.text.x =
    element_text(
      size = 10, angle = 0, hjust = .5, vjust = 1.5, face = "plain"
    ),
  axis.text.y =
    element_text(
      size = 10, angle = 0, hjust = 1, vjust = 0, face = "plain"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.title =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 10),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 12),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r include = FALSE}
# /*===========================================================
#' # Prepare data and results for writing
# /*===========================================================
#* source all the functions in the Functions folder
fs::dir_ls(here("Codes", "Functions"), full.names = TRUE) %>%
  lapply(., function(x) source(x))

#* read the results rds file
results <-
  here("Results", "pi_data.rds") %>%
  readRDS() %>%
  .[, type := ifelse(transfer == 0, "GWR-R", "GWR-T")] %>%
  .[, bias := pi_diff_est - pi_diff]

pi_dif_gwr <-
  results[, .(pi_dif = mean(pi_diff)), by = type] %>%
  .[, diff(pi_dif)]

#* create figures
source(here("Codes", "3_make_figures_publication.R"))
```

**Abstract**: Geographically-weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to generate site-specific input rate recommendations using data generated by on-farm precision experimentation (OFPE). The key objectives of this paper are two-fold. First, it evaluates two sets of functional-form specifications made by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that one set of functional-form specification performs significantly better than the other. Second, we investigate the possibility of bias when GWR is used to estimate the value of site-specific vs. uniform-rate input management. We show that GWR suffers significant positive bias, which can lead to great over-estimation of the value of applying GWR methods. Our study is the first to systematically examine the statistical and economic properties of the use of GWR in OPFE settings. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management.

**Keywords**: on-farm precision experimentation, site-specific input management, geographically-weighted regression, economically optimal nitrogen rate

**Acknowledgement**: This research was supported by a USDA-NIFA-AFRI Food Security Program Coordinated Agricultural Project, titled “Using Precision Technology in On-farm Field Trials to Enable Data-Intensive Fertilizer Management,” (Accession Number 2016-68004-24769), and also by the a USDA-NRCS Conservation Innovation Grant from the On-farm Trials Program, titled “Improving the Economic and Ecological Sustainability of US Crop Production through On-Farm Precision Experimentation” (Award Number NR213A7500013G021).

`r run_pagebreak()`

# Introduction

In recent years, On-farm Precision Experimentation (OFPE) has gained traction with both researchers and practitioners as a means of generating high-resolution spatial data to inform site-specific crop input management decisions [@bullock2019data; @lacoste2021farm]. Accompanying interest is surging among researchers in identifying efficient statistical methods to estimate site-specific yield response functions, which have examined various methods, ranging from Geographically Weighted Regression (GWR) [@rakshit2020; @evans2020assessment; @trevisan2021spatial] to machine learning methods including random forest [@krause2020random] and convolutional neural network [@barbosa2020modeling] approaches. Among these methods applied to OFPE data, GWR is distinct in that it does not require potentially costly collection of site-specific field and soil information in estimating site-specific yield response functions.

On any given field, it is typical for U.S. Midwestern row crop farmers to pay for 1-ha grid soil sampling and chemical analysis every three or four years [@morris2018strengths]. The data is often used to derive fertilizer application recommendations, for example by applying comparatively simple formulas provided by university extension services (e.g., Culman, et al. 2020). But the expense of soil sampling and analysis is driving development of sensor technologies that gather data on site-specific field and crop characteristics or proxies of those characteristics. For example, Trimble Agriculture's GreenSeeker and AgLeader's OptRx® systems use optical sensors mounted on fertilizer application booms to digitally record vegetative index imagery as farm machinery is driven through fields during early crop growth stages. GreenSeeker then basically uses Oklahoma State University's Sensor-based N Rate Calculator to convert those data into recommended nitrogen fertilizer (N) application rate maps ($R_xs$). Veris Technologies manufactures sensor equipment that can be towed through fields to gather apparent electroconductivity (EC) data, which can be used to infer properties of soil content. The "decision tool" software FieldRevealTM allows users to upload EC data, elevation data, and vegetative index data, which FieldRevealTM's professional agronomists analyze to provide $R_x$s to farmers. FieldApex creates N $R_x$s using foliar vegetative index data from satellite imagery, and SlantRange uses drone imagery for similar purposes.

However, deploying sophisticated sensors to gather field characteristics data is also costly, and the effects of their generation and use on farm profitability remains in question [@gardner2021economic]. Recognizing the significant costs of obtaining useful, high-resolution soil and field properties data, @trevisan2021spatial demonstrated a method of providing site specific management recommendations without such data. The study estimated site-specific yield response functions by applying GWR techniques on data from OFPEs [@bullock2019data]. @trevisan2021spatial reported dramatic results, estimating an average potential gain of \$65 ha$^{-1}$ from applying seed and N at optimal uniform rates (instead of at the rates the farmers would have chosen if not participating in the research), and an additional potential gain of \$58 ha$^{-1}$ when changing from optimal uniform application rates to optimal site-specific application rates. @rakshit2020 also applied GWR to on-farm experiment data. However, while they did estimate site-specific yield response functions, they did not derive site-specific economically optimal nitrogen rates based on them, and thus did not conduct economic analysis unlike @trevisan2021spatial. @evans2020assessment explored using GWR with a linear model specification to detect low- and high-response areas, but like @rakshit2020 also did not estimate site-specific economically optimal nitrogen rates.

This study has two key objectives. The first is to determine which of the model specifications used by @rakshit2020 and @trevisan2021spatial result in more profitable $R_xs$. While both @rakshit2020 and @trevisan2021spatial proposed using GWR to estimate site-specific yield response functions, they use different model specifications. Both assume a quadratic response of yield to nitrogen rate ($N$). However, while @trevisan2021spatial estimate the coefficient on $N^2$ as a global parameter, @rakshit2020 estimate it locally for each point. At first, it appears that the specification used by @rakshit2020 is superior since it is more flexible than the specification used by @trevisan2021spatial. However, the flexibility can be a disadvantage rather than an advantage because the estimation of EONR is highly sensitive to the coefficient estimates on $N^2$, and GWR often provides inaccurate local estimation of that coefficient.

Second, we will examine the degree and direction of bias in the estimation of the value of GWR-based variable rate application (VRA) and question the extremely high estimates of that value reported by @trevisan2021spatial. In general, researchers take the following steps to estimate the value of VRA based using OFPE data [e.g., @swinton1998evaluating; @bullock2002adding; @anselin2004spatial; @liu2006site]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIR) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the assumption made on step (3), which is that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provide an example. At step (3), they used their estimated GWR as the true yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true yield response function of any real field. Use of these procedures assumes accurate estimation of yield response functions, which when untrue can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions. We suspect that the high value of VRA estimated by @trevisan2021spatial displays this bias.

In the following, we pursue the two objectives just described above by conducting Monte Carlo simulations in which we first generate and then analyze stylized OFPE data. We cannot test which model specification results in more profitable site-specific nitrogen rate recommendations using real data because it is impossible to observe the true yield response functions. For the same reason, we cannot test whether @trevisan2021spatial over-estimates the actual value of VRA, either. But in the Monte Carlo simulations we specify yield response functions and generate stylized data from those functions, which enables us to quantify in the context of the simulations the value of GWR and the bias in GWR-based estimation of the value of VRA. Monte Carlo simulations are commonly used in this way to gain insights into the statistical properties of trial designs and estimations methods [@mckinion2001analysis; @gebbers2010application; @bullock2020value; @alesso2021design].

One of the principal findings of this article is that the model specification used by @trevisan2021spatial creates more valuable information than does the model specification used by @rakshit2020. This is primarily due to GWR's poor local estimation of the coefficients on $N^2$. Since the coefficients on $N^2$ is highly influential in determining site-specific EONR, inaccuracy in its estimation can result in significant economic loss. We also show that even under the method used by @trevisan2021spatial, the benefit of VRA over URA is rather small, averaging only about \$5 ha$^{-1}$; we conclude that both @rakshit2020 and @trevisan2021spatial significantly over-estimated that value. Our results suggest that the prospects of using GWR to generate site-specific input recommendations are bleaker than stated in @trevisan2021spatial.

# Materials and Methods

## Experimental Field Layout

Figure \@ref(fig:field-layout) shows the field layout and defines the spatial units of the simulated on-farm precision experiments in this study. The basic spatial unit of the field is the 6m $\times$ 6m "cell", which is the resolution at which the field characteristics data (i.e., parameters of the "true" yield response function) were allowed to vary. It was assumed, however, that yields could not be monitored at such a high resolution. Rather, the mechanical harvester was assumed to be 18m wide, and, in cases in which the N rate was unchanged, in need of harvesting on a run of no less than 12m to provide accurate yield data. Therefore, the yield data's resolution was 18m $\times$ 12m (3 cells $\times$ 2 cells, called a "subplot"). The N rate unit consisted of five "subplots", that is, 18m $\times$ 72m (3 cells $\times$ 12 cells, called a "plot") to accommodate the variable-rate applicator's application resolution. A 12m "transition zone" was placed between two plots (that is, a 6m "transition zone" on each end of a plot), so that when the experimental N rate changed, the harvester had 12 m to "clear" and begin to record accurately on a plot with an N rate different from its predecessor's.

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6-m-wide headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12-m-wide "transition zones" were not included in the data analyzed. In total there were 1440 subplots available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $k \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet (j, k) is used to identify individual cells.

## True Cell-specific Yield Response Function

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. The quadratic, linear-plateau, quadratic-plateau, and Mitscherlich‐Baule are commonly used functional forms. In this study, each cell (j, k)'s "true" yield response function is assumed to follow the quadratic-plateau functional form:

$$
\begin{equation}
y_{j,k} = f_{j, k} (N) = f(N, c_{j,k}) = 
  \begin{cases}
  \alpha_{j,k} + \beta_{j,k} N + \gamma_{j,k} N^2 + \varepsilon_{j,k} \, ,& N < \tau_{j,k} \\
  \alpha_{j,k} + \beta_{j,k} \tau_{j,k} + \gamma_{j,k} \tau_{j,k}^2 + \varepsilon_{j,k} \, ,& N \ge \tau_{j,k}
  \end{cases}
\end{equation}
$$

where $c_{j,k} = (\alpha_{j,k},\beta_{j,k},\gamma_{j,k},\tau_{j,k})$ is the vector of yield response function parameters. $\tau_{j,k}$ can be interpreted as the critical value of N rate at which cell $(j, k)$'s yield reaches the plateau level (i.e., where yield no longer increases with additional N fertilizer). $\varepsilon_{j, k}$ is a stochastic error term. The quadratic-plateau functional form implies a diminishing marginal product of N and a yield plateau, which are desirable properties for a yield response function [@frank1990comparison]. It is a widely observed corn nitrogen response pattern by existing agronomic studies. Note that the cell-specific yield response function $f_{j, k} (N)$ is the reduced form of the meta response function $f$ when the characteristics vector takes on the value $c_{j, k}$. In other words, all the soil or field properties that could affect yield response to N are embedded in the simulated parameters.

The spatial distributions of cell-level parameters $\alpha_{j, k}$, $\beta_{j, k}$, $\gamma_{j,k}$, $\tau_{j, k}$, and the error term $\varepsilon_{j, k}$ were simulated by generating spatially correlated random fields using unconditional Gaussian geostatistical simulation based on the specified variograms [@dietrich1996fast, @wood1994simulation], and the simulation processes were conducted using the R package gstat [@graler16]. Specifically, the model is spherical, the nugget variance is zero, and the range is 600 meters. The nugget value of zero means that infinitesimally close observations have a zero covariance (i.e., they are identical). The range of 600m means that observations separated by more than 600m are not spatially correlated. A higher range value indicates increased spatial dependence, meaning less abrupt differences are observed between the production function parameters of neighboring cells. The mean and sill variance of each parameter were specified to result in a realistic mean yield of $11,416$ kg ha$^{-1}$, and such that the cells' true economically optimal N rate ranged from $113$ kg ha$^{-1}$ to $273$ kg ha$^{-1}$ among all the simulations, which most real-world N rate recommendations fall into (e.g., @sawyer2018). Note that the range of the optimal N rates of individual rounds of simulations are narrower. In defining the magnitude of the error term, we utilized the experimental data generated through eight on-farm randomized N field trials conducted by the DIFM project in 2018 by the Data-Intensive Farm Management Project (DIFM), [@bullock2019data]. Using the data from each of the eight actual trials, separate regressions of yield on N were run, and standard deviations of the residuals were calculated. The mean value of the eight standard deviations was $1370$ kg ha$^{-1}$, which was the value of sill variances assigned to the simulations' error terms.

## Trial Design

Figures \@ref(fig:field-layout) and \@ref(fig:field-N-design) illustrate the Latin square trial design schema assumed in each simulation. The field was partitioned into a 4-block $\times$ 2-block grid, in which each block was partitioned into a 6-plot $\times$ 6-plot grid with each of six N rates assigned to exactly one plot in each row and each column of the block. The design ensured that no adjacent N plots were assigned the same N rate. Furthermore, the order of N rates in each row of every block was specified to avoid relatively large changes in N rates (due the technological limitations of N application equipment commonly witnessed in real-world OFPE).

Letting $p$ represent the crop price and $w$ represent the price of nitrogen fertilizer, in each simulation the "true" economically optimal N rate for each cell $(j, k)$ was calculated as

$$
N^*_{j,k} \equiv argmax_N \Big[p\cdot f_{j,k}(N)-w\cdot N\Big].
$$

In each simulation, six trial N rates were assigned by calculating the 0%, 20%, 40%, 60%, 80%, and 100% quantiles of the simulation's true cell-specific optimal N rates. The means of these trial N rates across simulations were approximately, {80, 130, 154, 184, 219, and 270} kg ha$^{-1}$.

## Yield Data Generation

In each simulation, for each cell $(j, k)$, the simulation's true yield $y_{j, k}$ was generated based on the cell's assigned application rate $N_{j,k}$, true response parameters $c_{j,k} = (\alpha_{j,k}, \beta_{j,k}, \gamma_{j,k}, \tau_{j,k})$, and error term $\varepsilon\_{j,k}$. This procedure generated a field-level, cell-specific data set, $\{(y_{j, k},N_{j, k} ):j=1,\dots,1440;k=1,\dots, 6 \}$. To maintain consistency with real-world OFPEs, it was assumed that harvesting equipment could only record yields accurately at an 18m $\times$ 12m (subplot) resolution. Since every cell in a plot received the same N application rate, then so did every cell within a subplot, so $N_{j,1} = N_{j,2}= \dots = N_{j,6} = N_j$ for each subplot $j$. Letting $\mu_j$ denote the mean of the yields of the cells in subplot $j$, the analysis of each round's data was based on 1440 $(\mu_j,N_j )$ observations.

## Estimation of Yield Response Function and EONR

Three approaches to estimate yield response functions and their associated EONRs were used. The first approach used the shape-constrained additive model (SCAM) to estimate a single yield response function for the entire field, and to estimate and economically optimal uniform nitrogen application rate for the field. The other two approaches both used GWR, but assumed different specifications in modeling the impact of nitrogen on yield. The first GWR approach (GWR-R) followed the specification used by @rakshit2020, and the second GWR approach (GWR-T) followed the specification used by @trevisan2021spatial. Further details of each approach are presented below. SCAM and GWR estimations were performed using $R$'s scam [@scam-R] and GWmodel [@gwmodel-R] packages.

### Spatially Uniform Management after Non-Spatial Regression Analysis of OFPE Data

In the lower benchmark situation, a researcher was assumed to run OFPEs to gather data, then to estimate a single "full-field" yield response function using the SCAM non-parametric regression framework, which is a spline-based series estimation that can impose a shape-constraints on the estimated function [@pya2015shape]. In using SCAM, the yield response function was restricted to be monotonically increasing and concave with respect to the N application rate, which are traits well substantiated by agronomic research. When the chosen shape constraints are correct, SCAM is more statistically efficient than a fully non-parametric regression [@pya2015shape]. The estimated model is represented mathematically as follows,

$$y_j = \alpha + \sum_{l=1}^l \beta_l g_l(N_j)+ \varepsilon_j$$

where $g_l(\cdot)$ is the $l$th spline basis and $\beta_l$ is its coefficient to be estimated. The $\beta_l$s were constrained such that the resulting relationship between $y_j$ and $N_j$ is such that $y_j$ monotonically increases as $N_j$ increases, but at a diminishing rate.

Let $\hat{f}_{scam}(\cdot)$ represent the estimated whole-yield yield response function ($\sum_{k=1}^K \hat{\beta}_k g_k(\cdot)$). Given the information derived from the SCAM approach, the estimate of the economically optimal uniform N rate is,

$$\hat{N}_{scam}^* \equiv argmax_N⁡ \big[p\cdot \hat{f}_{scam} (N)-w\cdot N\big]$$

### GWR-Rakshit (GWR-R)

In the GWR-R approach, yield response was modeled as subplot-specific quadratic functions of the nitrogen rate, which for each subplot $j \in {1, 2, \dots, 1440}$ was specified as,

$$y_j= \alpha_j+ \beta_j N_j + \eta_j N_j^2+ \epsilon_j$$

The GWR-R estimation produced subplot-specific estimates of coefficients $\hat{\alpha}_j$, $\hat{\beta}_j$, and $\hat{\eta}_j$, and so of the yield response function $\hat{f}_{gwr-r,j}(N)$ for $j = 1, 2, \dots, 1440$. The estimated economically optimal N application rate for subplot $j$ was calculated by solving the subplot's profit maximization problem:

$$\hat{N}_{gwr-r,j}^{*} \equiv argmax_N⁡[p\hat{f}_{gwr-r, j} (N) -wN]$$.

### GWR-Trevisan (GWR-T)

In the GWR-T approach, unlike GWR-Rakshit, the coefficient on $N^2$ was estimated as a global parameter while the coefficient on $N$ was estimated subplot-specifically. For each subplot $j \in {1, 2, \dots, 1440}$, the assumed yield response was,

$$y_j= \alpha_j+ \beta N_j + \eta N_j^2+ \epsilon_j$$.

where $\eta$ was assumed invariant among subplots. Because the coefficients on some variables were estimated as global parameters, the model used can be classified as a Mixed GWR model.

Let $\hat{f}_{gwr-t,j}(N)$ for $j = 1, 2, \dots, 1440$ denote the estimated yield response function for subplot $j$. The estimated economically optimal N application rate for subplot $j$ was then,

$$\hat{N}^*_{gwr-t,j} \equiv argmax_N⁡[p\hat{f}_{gwr-t,j} (N) -wN]$$.

## The value of GWR-based VRA over scam-based URA

To estimate the bias in the estimation of the value of GWR-based VRA, the true per-ha net revenues ("profits") of the SCAM-based URA and GWR-based VRA were calculated by plugging in the estimated optimal nitrogen rates into the "true" yield response functions. Profits from the SCAM approach were

$$\Pi_{scam}^* = 1/8640 \sum_{j=1}^{1440} \sum_{k=1}^6 \big[pf_{j, k} (\hat{N}_{scam}^* )-w\hat{N}_{scam}^*\big]$$

where $f_{j, k}(\cdot)$ denotes the true yield response function at cell $(j, k)$.

For $m = R, T$, the true per-ha net revenues ("profits") resulting from applying the profit-maximizing cell-specific N rate plan estimated by GWR-m were,

$$\Pi_{gwr-m}=1/8640 \sum_{j=1}^{1440} \sum_{k=1}^6 [pf_{j, k} (\hat{N}_{gwr-m, j}^*)-w\hat{N}_{gwr-m,j}^*]$$

where $f_{j, k} (\hat{N}_{gwr-j}^*)$ denote the true yield generated at cell $(j, k)$ with subplot $l$'s N rate.

The true value of the GWR-based VRA compared to the SCAM-based URA is defined by the difference of their partial profits:

$\Delta_{gwr-m} = \Pi_{gwr-m} - \Pi_{scam}$

To find the bias in estimating the value of VRA, the "estimated" profits of the SCAM-based URA and GWR-based VRA were calculated by plugging in the estimated optimal nitrogen rates into the cell-specific yield response functions estimated by GWR. Estimated per-ha net revenues for SCAM-based URA and GWR-based VRA are found as follows:

$$\hat{\Pi}_{scam}^* = 1/8640 \sum_{j=1}^{1440} \sum_{k=1}^6 \big[p\hat{f}^{gwr}_{j, k} (\hat{N}_{scam}^* )-w\hat{N}_{scam}^*\big]$$

$$\hat{\Pi}_{gwr-m}^*=1/8640 \sum_{j=1}^{1440} \sum_{k=1}^6 [p\hat{f}^{gwr-m}_{j, k} (\hat{N}_{gwr-m}^{j*} )-w\hat{N}_{gwr-m}^{j*}]$$

The estimated value of the GWR-based VRA compared to the SCAM-based URA is defined as

$$\hat{\Delta}_{gwr-m} = \hat{\Pi}_{gwr-m}^* - \hat{\Pi}_{scam}^*$$

Finally, the bias in the estimation of the value of GWR-based VRA was calculated as follows:

$$bias = \hat{\Delta}_{gwr-m} - \Delta_{gwr-m}$$

# Results

## Comparing the values of GWR-based VRA and SCAM-based URA

Figure \@ref(fig:pi-dif-dist) shows the distribution of $\hat{\Delta}_{gwr-m}$, profit difference between the GWR-based VRA and the SCAM-based URA. The distribution of the value of VRA from GWR-R is centered around 0, indicating that VRA based on GWR-R is on average no more profitable than SCAM-based URA. This result comes from GWR-R's highly inaccurate estimation of the subplot-level $N$ and $N^2$ coefficients, which led to unreliable subplot-specific nitrogen rate recommendations. Figure \@ref(fig:true-vs-estimated-coef-gwr-r) illustrates the accuracy of GWR-R in estimating the $N$ and $N^2$ coefficients in one simulation round. Estimates of the $N^2$ coefficients are positive for some subplots, indicating an increasing marginal product of the nitrogen rate, which contradicts a well-established agronomic property. The result is consistent with the bottom panel of Figure 12 in @rakshit2020. Figure \@ref(fig:true-vs-estimated-optn-gwr-r) plots the estimated cell-specific optimal N against the true optimal N for the same simulation round, and again shows that GWR-R can be highly unreliable, and consequently that basing variable rate management on GWR-R may be no better than taking a simple URA approach. Basically, not enough information is available locally to estimate reliably the coefficients on both $N$ and $N^2$ by subplot.

The GWR-T approach, however, performed better than SCAM-URA in majority of cases, with some cases showing values-added exceeding \$10 ha$^{-1}$. This may seem surprising as the model estimated in the GWR-T approach is more restrictive than that in the GWR-R approach because the coefficient on $N^2$ is assumed to be global. However, unlike GWR-R, this restriction helps avoid estimating positive subplot-level coefficients on $N^2$. This result implies that the value of avoiding highly unreliable estimates of coefficients on $N^2$ subplot-specifically was greater than loss of local flexibility in the coefficients on $N^2$.

## Bias in the estimation of the value of GWR-based VRA

Next we examine the bias in the "estimated" value of GWR-based VRA that researchers would encounter using real data. Figure \@ref(fig:bias-est-pi) shows the distribution of the bias in "estimated" profit between the GWR-based VRA and SCAM-based URA. Unlike the true difference in profit, VRA based on GWR-R appears to perform overwhelmingly better, with its average profit about \$`r round(pi_dif_gwr, digits = 2)` ha$^{-1}$ greater than that of the SCAM-based URA. That is, researchers would tend to significantly over-estimate the value of VRA based on GWR-R. When the GWR estimates are considered to define true yield response, the inaccuracy of GWR-R generates over-estimation of the heterogeneity of yield response within the field, which can lead to significant over-estimation of the value of GWR-R itself. Under GWR-T, the bias is more than \$10 ha$^{-1}$ for `r results[type == "GWR-T", mean(abs(bias) > 10)] * 100` % of the simulation cases: under GWR-R, that bias is greater than \$10 ha$^{-1}$ in `r results[type == "GWR-R", mean(abs(bias) > 10)] * 100` % of the simulated cases. However, as can be seen in the bottom panel of Figure \@ref(fig:bias-est-pi), the distribution of bias from GWR-T has a fatter tail than GWR-R with more mass on higher bias values. Bias is more than \$50 ha$^{-1}$ for `r results[type == "GWR-T", mean(abs(bias) > 50)] * 100` % of the simulation cases for GWR-T as opposed to `r results[type == "GWR-R", mean(abs(bias) > 50)] * 100` % for GWR-R.

Figure \@ref(fig:why-bias-many) illustrates how over-estimations can occur. It shows yield response functions estimated by GWR-T and estimated yields associated with the GWR-based EONR rates and SCAM-based EONR rates for 50 randomly chosen cells in a single simulation. Since GWR-based EONRs were obtained assuming the yield response functions estimated by GWR, by construction they out-perform better the SCAM-based estimates in every cell. For many cells, the SCAM-based estimated EONR is significantly higher than the actual yield-maximizing N rate, resulting in a yield far lower because of the quadratic functional form assumption. Figure \@ref(fig:why-bias-single) focuses on a single cell and compares estimated and true yields between GWR and SCAM (panel (a)) and estimated and true profits between GWR and SCAM (panel (b)). The cell's true yield response functions plateaus at a nitrogen rate of about $140$ kg ha$^{-1}$, but the GWR-estimated yield response function suggests that yield should peak at a nitrogen rate of about $190$ kg ha$^{-1}$ and then yield declines afterward. In reality, the yield differential between GWR and SCAM is 0, while the estimated yield differential is about $0.25$ ton ha$^{-1}$, leading to the over-estimation of the value of GWR in panel (b). Based on the true profit response curve, applying the GWR-based recommendation is about \$`r round(- true_yield_data[type != "True", diff(profit)], digits = 0)` ha$^{-1}$ more profitable than applying the SCAM-based recommendation. This profit differential comes soley from the difference in nitrogen cost since GWR and SCAM yields are equal. However, if calculations are based on the cell's GWR-based estimation of the response function, profits from the GWR-based recommendation are measured to be about \$`r round(- point_data_gwr[, diff(profit)], digits = 0)` ha$^{-1}$ greater than from the SCAM-based recommendation. Beyond the difference in nitrogen cost, the falsely estimated yield boost by GWR contributes to the profit differential, thus significantly over-estimating the actual value of GWR.

# Discussions

Our results on the relative performance of GWR-T and GWR-R show clearly that GWR-T performs better than GWR-R, and GWR-T should be used when researchers employ a GWR-based approach to find site-specific input recommendations. These findings, however, have important implications beyond just GWR-R and GWR-T. An obvious extension of the parametric GWR approaches would be a semi-parametric or non-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). However, such a method is likely to encounter a very similar problem to that faced by GWR-R: insufficient information available locally to estimate reliably the shape of yield response curve locally.

It is important to be aware that researchers need to be extremely cautious about their estimated value of VRA based on GWR whether they follow GWR-R or GWR-T. Our results show that it is likely that they will over-estimate their true values and the bias can be misleadingly high at the non-negligible probability. Extremely high values of VRA reported in @trevisan2021spatial are likely to have reflected this bias. Indeed, the bias problem we raised here is likely to be observed for other statistical methods because the bias comes from the assumption that the _estimated_ yield response functions are true. Whatever the underlying statistical method for yield response functions may be, researchers are likely to overestimate the true value of their methods. It would be an interesting future research to investigate the bias in other major statistical methods that are commonly used in the literature.

# Conclusions

This articles examined the value of GWR-based VRA and the bias in its estimation in the context of OFPE. We showed that restricting the coefficient on $N^2$ to a global parameter as done in @trevisan2021spatial leads to site-specific nitrogen rate recommendations with higher profits compared to the specification used by @rakshit2020, which allows the coefficients on $N^2$ to vary locally. However, we also found that both specifications can result in significant over-estimation of the true value of GWR-based site-specific nitrogen rate recommendations. It can be concluded that researchers need to be cautious about promoting the GWR-based approach.

A fruitful extension of GWR-based approach would be to employ more flexible yield reponse model specifications. This study has shown the assumed quadratic functional form to be the main culprit of the significant over-estimation of the value of GWR-based VRA. Applying GWR with a quadratic-plateau or linear-plateau specifications might substantially alleviate the reported bias problem.

`r run_pagebreak()`

# Figures

```{r fig.id = "field-layout", fig.cap = "Simulated field layout with spatial unit definitions", fig.dim = c(6, 6)}
g_layout
```

```{r fig.id = "field-N-design", fig.cap = "Experiment design of nitrogen (N) rates"}
g_exp
```

```{r, fig.id = "pi-dif-dist", fig.cap = "The value of GWR-based VRA over SCAM-based URA for GWR-R and GWR-T"}
g_value
```

```{r fig.id = "true-vs-estimated-coef-gwr-r", fig.cap = "Comparison of Estimated and True Coefficients"}
g_comp_coef
```

```{r fig.id = "true-vs-estimated-optn-gwr-r", fig.cap = "Comparison of Estimated and True EONR"}
g_comp_eonr
```

```{r fig.id = "bias-est-pi", fig.cap = "Bias in the estimation of the value of GWR-based VRA over SCAM-based URA for GWR-R and GWR-T"}
g_bias
```

```{r fig.id = "why-bias-many", fig.cap = "The cause of significant over-estimation of the value of GWR-based VRA"}
g_why_bias_many
```

```{r fig.id = "why-bias-single", fig.cap = "An illustration of over-estimation of the value of GWR-based VRA over SCAM-based URA",fig.dim = c(6, 7)}
g_why_bias_single
```

`r run_pagebreak()`

# References

<br />

<div id="refs"></div>