### 2022-02-01 Writing/manuscript_gwr.rmd
rmd template added

---
title: "New"
author:
  - Taro Mieno^[University of Nebraska Lincoln, tmieno2@unl.edu], Xiaofei Li^[Mississippi State University, xiaofei.li@msstate.edu], David S. Bullock^[University of Illinois, dsbulloc@illinois.edu]
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: true
    reference_docx: "word_template.docx"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ": "
bibliography: PA.bib
csl: field-crops-research.csl
abstract: ""
---

```{r echo = F, cache = F}
suppressMessages(library(knitr))
suppressMessages(library(here))
suppressMessages(library(officedown))
suppressMessages(library(officer))

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = F,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(lfe)
library(modelsummary)
library(patchwork)
```

```{r figure_setup, cache = F}
theme_update(
  axis.title.x =
    element_text(
      size = 12, angle = 0, hjust = .5, vjust = -0.3, face = "plain"
    ),
  axis.title.y =
    element_text(
      size = 12, angle = 90, hjust = .5, vjust = .9, face = "plain"
    ),
  axis.text.x =
    element_text(
      size = 10, angle = 0, hjust = .5, vjust = 1.5, face = "plain"
    ),
  axis.text.y =
    element_text(
      size = 10, angle = 0, hjust = 1, vjust = 0, face = "plain"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.title =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 10),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 12),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r include = FALSE}
# /*===========================================================
#' # Prepare data and results for writing
# /*===========================================================
#* source all the functions in the Functions folder
fs::dir_ls(here("Codes", "Functions"), full.names = TRUE) %>%
  lapply(., function(x) source(x))

#* read the results rds file
results <-
  here("Results", "pi_data.rds") %>%
  readRDS() %>%
  .[, type := ifelse(transfer == 0, "GWR-R", "GWR-T")] %>%
  .[, bias := pi_diff_est - pi_diff]

pi_dif_gwr <-
  results[, .(pi_dif = mean(pi_diff)), by = type] %>%
  .[, diff(pi_dif)]

#* create figures
source(here("Codes", "3_make_figures_publication.R"))
```

**Abstract**:

**Keywords**:

**Acknowledgement**: This research was supported by a USDA-NIFA-AFRI Food Security Program Coordinated Agricultural Project, titled “Using Precision Technology in On-farm Field Trials to Enable Data-Intensive Fertilizer Management,” (Accession Number 2016-68004-24769), and also by the a USDA-NRCS Conservation Innovation Grant from the On-farm Trials Program, titled “Improving the Economic and Ecological Sustainability of US Crop Production through On-Farm Precision Experimentation” (Award Number NR213A7500013G021).

`r run_pagebreak()`

# Introduction

# Methods

# Results

# Discussions

`r run_pagebreak()`

# Tables

`r run_pagebreak()`

# Figures

`r run_pagebreak()`

# References

<br />

<div id="refs"></div>