### 2022-02-03 Codes/A0_prepare_data_for_sharing.rmd
program to prepare data for others created

# Objective

+ Create a dataset to share with Paul

# Set up

## Packages
```{r }
library(sf)
library(data.table)
library(tidyverse)
library(here)
```

# Prepare Data

```{r }
#* Read field data
field_data <-
  readRDS(here("Shared/Data/field_data.rds")) %>%
  pull(field_sf) %>%
  .[[1]]

saveRDS(field_data, here("Shared/Data/ToShare/field_sf.rds"))

field_with_design <- readRDS(here("Shared/Data/field_with_design.rds")) %>% .[sce_i, ]

x_vars <-
  c(
    "b0_1", "b0_2",
    "Nk_2_1", "Nk_2_2", "Nk_1_1",
    # "Nk_2_1", "Nk_2_2", "Nk_1_1", "Nk_1_2",
    "plateau_2_1", "plateau_2_2", "plateau_1_1",
    # "plateau_2_1", "plateau_2_2", "plateau_1_1", "plateau_1_2",
    "theta_plateau_1", "theta_plateau_2",
    "theta_Nk_1", "theta_Nk_2",
    "theta_b0_1", "theta_b0_2"
  )

#* Read the simulated data
sim_data <-
  field_with_design$data_file_name %>%
  readRDS() %>%
  .$reg_data %>%
  .[[1]] %>%
  .[, N_levels := NULL] %>%
  unnest() %>%
  data.table() %>%
  .[, c("sim", "aunit_id", x_vars), with = FALSE] %>%
  setnames(
    names(.),
    c(
      "sim",
      "aunit_id",
      paste0("par", 1:(length(names(.)) - 2))
    )
  )

saveRDS(sim_data, here("Shared/Data/ToShare/sim_data.rds"))
```