### 2022-02-04 Codes/A1_RF_sensitivity_check.rmd
block_id added to regression data for clustering

# Objective

+ Check how sensitive RF is to
  + the number of variables included
  + tuning parameters

# Conclusion

+ default parametrers are really good
+ tuning parameters via cross-validation does not really help

# Set up

## Packages
```{r }
library(sf)
library(data.table)
library(grf)
library(tidyverse)
library(here)
```

```{r }
fs::dir_ls(here("GitControlled", "Codes", "Functions", "R"), full.names = TRUE) %>%
  purrr::map(~ source(.))

#* field
sce_i <- 3
```

# Prepare Data

```{r }
#* field sf data
field_data <-
  readRDS(here("Shared/Data/field_data.rds")) %>%
  pull(field_sf) %>%
  #* largest field
  .[[sce_i]]

#* field parameters
field_parameters <-
  readRDS(here("Shared/Data/field_parameters.rds")) %>%
  .[sce_i, ]

#* number of simulation cases
nsim <- field_parameters$field_pars[[1]][, max(sim)]

#* Read the simulated data
field_with_design <-
  readRDS(here("Shared/Data/field_with_design.rds")) %>%
  .[sce_i, ]

field_sf <- field_with_design$field_sf

sim_data <-
  field_with_design$data_file_name %>%
  readRDS()

#* define prices
pN <- sim_data$pN
pCorn <- sim_data$pCorn
```

## Define simulation function

```{r }
run_analysis <- function(sim_i, x_vars, models_data, pN, pCorn, Wls, nsim) {
  print(sim_i)

  # /*+++++++++++++++++++++++++++++++++++
  #' ## Prepare data
  # /*+++++++++++++++++++++++++++++++++++
  # /*+++++++++++++++++++++++++++++++++++
  #* extract the data for the sim_i
  train_data <- sim_data$reg_data[[1]][sim == sim_i, ]
  cv_data <- sim_data$reg_data[[1]][sim == ifelse(sim_i + 1 > nsim, 1, sim_i + 1), ]

  #* define parameters
  N_levels <- train_data$N_levels[[1]]
  reg_data <-
    train_data$data[[1]] %>%
    .[, x := (X - min(X)) / (max(X) - min(X))] %>%
    .[, y := (Y - min(Y)) / (max(Y) - min(Y))] %>%
    .[, xy := x * y] %>%
    .[, N2 := N^2]

  # reg_data[, N] %>% hist(breaks = 30)
  # ggplot(reg_data) +
  #   geom_point(aes(y = yield, x = N))

  results <-
    models_data %>%
    .[on == TRUE, ] %>%
    rowwise() %>%
    mutate(results = list(
      run_indiv_analysis(model, reg_data, cv_data, x_vars, pN, pCorn, N_levels, Wls)
    )) %>%
    mutate(results = list(
      mutate(results, model = model)
    )) %>%
    pull(results) %>%
    rbindlist()

  return(results)
}
```

## Define RF Analysis

```{r }
# /*===========================================================
#' # RF analysis
# /*===========================================================
# ! Run RF and do economic analysis

run_rf_analysis <- function(reg_data, cv_data, x_vars, pN, pCorn, N_levels, include_int = FALSE) {
  train_data <- copy(reg_data)
  test_data <- copy(cv_data$data[[1]])

  # /*+++++++++++++++++++++++++++++++++++
  #' ## Define X
  # /*+++++++++++++++++++++++++++++++++++
  if (include_int == FALSE) {
    all_X_vars <- x_vars
  } else {
    int_var_names <- paste0("N_", x_vars)
    int_var_exp <- paste0("= N * ", x_vars)
    int_var_gen_exp <- paste0(int_var_names, int_var_exp, collapse = ",")

    all_X_vars <- c(x_vars, int_var_names)

    eval(parse(text = paste("train_data[, `:=`(", int_var_gen_exp, ")]", sep = "")))
    eval(parse(text = paste("test_data[, `:=`(", int_var_gen_exp, ")]", sep = "")))
  }

  X <- train_data[, c("N", all_X_vars), with = FALSE]

  # /*+++++++++++++++++++++++++++++++++++
  #' ## Define Y
  # /*+++++++++++++++++++++++++++++++++++
  Y <- train_data[, yield]

  # /*+++++++++++++++++++++++++++++++++++
  #' ## Run RF
  # /*+++++++++++++++++++++++++++++++++++
  RF_temp <-
    grf::regression_forest(
      X = X,
      Y = Y,
      num.trees = 2000,
      # min.node.size = 10
      # tune.parameters = "all",,
      # tune.parameters = c("min.node.size", "sample.fraction", "mtry"),
      # tune.num.trees = 100,
      # tune.num.reps = 100
    )

  #************************   RMSE   ***********************
  e_hat_train <-
    train_data %>%
    .[, y_hat := predict(RF_temp, newdata = .[, c("N", all_X_vars), with = FALSE])] %>%
    .[, e_hat := yield - y_hat] %>%
    .[, .(aunit_id, e_hat)] %>%
    setnames("e_hat", "e_hat_train")

  e_hat_cv <-
    test_data %>%
    .[, y_hat := predict(RF_temp, newdata = .[, c("N", all_X_vars), with = FALSE])] %>%
    .[, e_hat := yield - y_hat] %>%
    .[, .(aunit_id, e_hat)] %>%
    setnames("e_hat", "e_hat_cv")
  #*********************************************************

  N_seq <-
    data.table(
      N = seq(min(train_data$N), max(train_data$N), by = 2)
    )

  rf_results <-
    test_data %>%
    .[, c("aunit_id", "yield", all_X_vars), with = FALSE] %>%
    expand_grid_df(., N_seq) %>%
    .[, yield_hat := predict(RF_temp, newdata = .[, c("N", all_X_vars), with = FALSE])] %>%
    .[, pi_hat := pCorn * yield_hat - pN * N] %>%
    .[, .SD[which.max(pi_hat)], by = aunit_id] %>%
    .[, opt_N_hat := N] %>%
    .[, .(aunit_id, opt_N_hat)] %>%
    #--- RMSE ---
    .[e_hat_train, on = "aunit_id"] %>%
    .[e_hat_cv, on = "aunit_id"] %>%
    .[, sim := cv_data$sim]

  return(rf_results)
}
```

# Analysis

```{r, include=FALSE}
models_data <-
  data.table(
    model = c(
      "ser_50", # spatial error, weights_50
      "ser_100", # spatial error, weights_100
      "ser_200", # spatial error, weights_200
      "ser_500", # spatial error, weights_500
      "lm", # linear quadratic
      "gwr_t", # gwr-trevisan
      "gwr_semi", # gwr-semiparametric
      "gwr_zone_scam", # scam by zone,
      "ma_cf", # multiarm-CF
      "brf", # boosted RF
      "rf", # random forest
      "dmlof_semi", # semiparmetric DML-OF
      "dmlof_quad", # semiparmetric DML-OF
      "drof" # DR-OF
    ),
    on = c(
      FALSE, # ser_50
      FALSE, # ser_100
      FALSE, # ser_200
      FALSE, # ser_500
      FALSE, # lm
      FALSE, # gwr_t
      FALSE, # gwr_semi
      FALSE, # gwr_zone_scam
      FALSE, # ma_cf
      FALSE, # brf
      TRUE, # rf
      FALSE, # dmlof_semi
      FALSE, # dmlof_quad
      FALSE # drof
    )
  )

x_vars <-
  c(
    "b0_1", "b0_2",
    "Nk_2_1", "Nk_2_2", "Nk_1_1", "Nk_1_2",
    "plateau_2_1", "plateau_2_2", "plateau_1_1", "plateau_1_2",
    "theta_plateau_1", "theta_plateau_2",
    "theta_Nk_1", "theta_Nk_2",
    "theta_b0_1", "theta_b0_2"
  )

x_vars <-
  c(
    "b0", "Nk", "plateau"
  )

sim_range <- c(1:2)

eonr_results <-
  lapply(
    sim_range,
    function(x) {
      run_analysis(
        sim_i = x,
        x_vars = x_vars,
        models_data = models_data,
        pN = pN,
        pCorn = pCorn,
        Wls = Wls,
        nsim = nsim
      )
    }
  ) %>%
  rbindlist()

est_result <-
  eonr_results %>%
  nest_by(model) %>%
  mutate(
    perform = list(
      field_pars[sim %in% (sim_range + 1), ] %>%
        data.table(field_sf)[, .(cell_id, aunit_id)][., on = "cell_id"] %>%
        data.table(data)[., , on = c("sim", "aunit_id")] %>%
        .[, y_hat := gen_yield_QP(b0, b1, b2, Nk, opt_N_hat)] %>%
        .[, profit := pCorn * y_hat - pN * opt_N_hat] %>%
        .[, profit := profit - profit_opt] %>%
        .[, .(
          profit = mean(profit, na.rm = TRUE),
          rmse_train = mean(e_hat_train^2, na.rm = TRUE) %>% sqrt(),
          rmse_cv = mean(e_hat_cv^2, na.rm = TRUE) %>% sqrt()
        ),
        by = sim
        ]
    )
  ) %>%
  mutate(
    field_col =
      field_with_design$field_col[sc_i],
    mean_profit =
      perform[, mean(profit, na.rm = TRUE)],
    rmse_train =
      perform[, mean(rmse_train, na.rm = TRUE)],
    rmse_cv =
      perform[, mean(rmse_cv, na.rm = TRUE)]
  )

est_result
```