### 2022-02-20 Writing/presentation.rmd
result 1st draft missing 2888 commits

---
title: "Pitfall of Prediction-orienterd Machine Learning Models and Economic Value of Statistical Efficiency in On-farm Precision Experimentation"
author:
  - Taro Mieno^[University of Nebraska Lincoln, tmieno2@unl.edu], Xiaofei Li^[Mississippi State University, xiaofei.li@msstate.edu], David S. Bullock^[University of Illinois, dsbulloc@illinois.edu]
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo = F, cache = F, include = F}
library(knitr)
library(here)

here::i_am("GitControlled/Writing/manuscript.rmd")

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = T,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F, include = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(officedown)
library(officer)
library(flextable)
library(stringr)
library(sf)
library(lfe)
library(modelsummary)
library(patchwork)
library(gridExtra)
```

```{r figure_setup, cache = F}
theme_update(
  axis.title.x =
    element_text(
      size = 12, angle = 0, hjust = .5, vjust = -0.3, face = "plain"
    ),
  axis.title.y =
    element_text(
      size = 12, angle = 90, hjust = .5, vjust = .9, face = "plain"
    ),
  axis.text.x =
    element_text(
      size = 10, angle = 0, hjust = .5, vjust = 1.5, face = "plain"
    ),
  axis.text.y =
    element_text(
      size = 10, angle = 0, hjust = 1, vjust = 0, face = "plain"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.title =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 10),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 12),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r include = FALSE}
# /*===========================================================
#' # Prepare data and results for writing
# /*===========================================================
#* source all the functions in the Functions folder
# fs::dir_ls(here("Codes", "Functions"), full.names = TRUE) %>%
#   lapply(., function(x) source(x))

# /*----------------------------------*/
#' ##  data preparation
# /*----------------------------------*/
#' load results data
est_data <-
  readRDS(here("Shared", "Results", "est_result_ls.rds")) %>%
  rbindlist() %>%
  dplyr::select(model, perform, field_col) %>%
  unnest(perform) %>%
  data.table()
#' rename columns
est_data <- est_data %>%
  .[, .(field_col, model, sim, profit, rmse_train, rmse_cv, rmse_eonr)] %>%
  #---field size---
  .[, field_size := paste0(round(field_col * 72 * 6^2 / 10000, 1), " ha")] %>%
  .[, field_size := factor(field_size, levels = c("9.3 ha", "18.7 ha", "37.3 ha"))] %>%
  #---model name---
  .[model == "brf", model := "BRF"] %>%
  .[model == "rf", model := "RF"] %>%
  .[model == "rf_perfect", model := "RF best"] %>%
  .[model == "lm", model := "LM"] %>%
  .[model == "ser_50", model := "SER"] %>%
  .[, model := factor(model, levels = c("RF", "RF best", "BRF", "LM", "SER"))]
#' mean and sd
mean_data <- est_data %>%
  .[, .(
    profit = mean(profit) %>% round(2),
    profit_sd = sd(profit) %>% round(2),
    rmse_yield = mean(rmse_cv) %>% round(2),
    rmse_yield_sd = sd(rmse_cv) %>% round(2),
    rmse_eonr = mean(rmse_eonr) %>% round(2),
    rmse_eonr_sd = sd(rmse_eonr) %>% round(2)
  ),
  by = c("field_size", "model")
  ] %>%
  .[order(field_size, model), ]
#' load illustration data (single simulation)
illus_data <-
  readRDS(here("Shared", "Results", "illus_data.rds"))

# /*----------------------------------*/
#' ##  create figures and tables
# /*----------------------------------*/
source(here("GitControlled", "Codes", "3_make_figures_tables.R"))
```

## Introduction

- Prediction-oriented Machine Learning (ML) methods swept the field of optimal input management in precision agriculture.
- Yield prediction is the ultimate goal.
- Big success.
- But unclear whether good yield prediction translates into farming profitability.

## Objective of This Study

- Compare the performance of prediction-oriented ML methods and traditional linear regression approaches in terms of the profitability of estimated site-specific nitrogen recommendation.
- Monte Carlo simulations
- On-farm precision experimental data

## Simulation Setting

```{r true-pars-b2, fig.cap="Simulated field", out.width="20%", out.height="20%", echo=FALSE, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "true_pars_b2.png"))
```

## True Yield Generation

Corn yield response to nitrogen (N) fertilizer at each location $i$
- Quadratic-plateau function:

$$
\begin{aligned}
y_{i} = 
  \begin{cases}
  \alpha_{i} + \beta_{i} N + \gamma_{i} N^2 + \varepsilon_{i}, & N < \tau_{i} \\
  \eta_{i}, & N \geq \tau_{i}
  \end{cases}
\end{aligned}
$$
- $\eta_i$: yield plateau
- $\tau_i$: the N rate where yield reaches plateau

## Field Characteristics Data

- Spatial variability of true parameters ($\alpha, \tau, \eta$) are generated by Gaussian process
- But producers do not observe them.
- Only observe the decomposition of the true parameters ($X_1$):

$$
\begin{aligned}
\alpha & = \alpha^1 + \alpha^2 \\
\tau & = min\big(min(\tau^{1,1}, \tau^{1,2}), min(\tau^{2,1}, \tau^{2,2})\big) \\
\eta & = min\big(min(\eta^{1,1}, \eta^{1,2}), min(\eta^{2,1}, \eta^{2,2})\big) 
\end{aligned}
$$
- and two nuisance covariates ($X_2$) that are correlated with $\alpha$, $\tau$, and $\eta$:

$$
\phi^{\alpha, 1}, \phi^{\alpha, 2}, \phi^{\tau, 1}, \phi^{\tau, 2}, \phi^{\eta, 1}, \phi^{\eta, 2}
$$

## On-Farm Nitrogen Experiment

- Five (5) nitrogen trial rates
- Latin Square experimental design

```{r N-trial-design, fig.cap="Simulated field", out.width="20%", out.height="20%", echo=FALSE, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "Nrate_Latin Square Fixed_blocked.png"))
```

## Estimation Models

Estimation of yield response functions and site-specific optimal nitrogen rates

- Machine Learning:
- Random Forest (RF)
- Random Forest with best possible data (true $\alpha$, $\tau$, and $\eta$)(RF$_{b}$)
- Boosted Regression Forest (BRF)

- Traditional Regression:
Linear Model OLS (LM)
Spatial Error Model (SER)

## Performance Measurements (1)

- Yield prediction
Root-mean-square error (RMSE) of out-of-sample predicted yield ($\hat{y}$):
$$
\begin{aligned}
RMSE_y = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (\hat{y}_i - y_i)^2}
\end{aligned}
$$
- Train data: One simulation round
- Test data: Another simulation round

## Performance Measurements (2)

- Economically optimal nitrogen rate (EONR) prediction:
Root-mean-square error (RMSE) of out-of-sample predicted EONR:

$$
\begin{aligned}
RMSE_{N} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (\hat{EONR}_i - EONR_i)^2}
\end{aligned}
$$

## Performance Measurements (3)

- Production profit: 
Generated by applying the estimated EONR:

$$
\begin{aligned}
\pi_i = p_{corn} f_i(\hat{EONR}_i) - p_N N_i - C
\end{aligned}
$$
- $f_i()$: true yield response function

## Results

- Machine learning models perform well in yield prediction,

- but poor in EONR prediction

- Traditional SER model generates the highest profit

## An Example Simulation

```{r single-simu-yield, fig.cap="Example simulation, yield.", fig.width=6, fig.height=7.5, echo=FALSE, cache=FALSE}
single_simu_yield
```

```{r single-simu-EONR, fig.cap="Example simulation, EONR", fig.width=6, fig.height=7.5, echo=FALSE, cache=FALSE}
single_simu_EONR
```

## 1,000 Simulations
```{r comb-boxplot-large, fig.cap="The performances of different models", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
comb_boxplot_large
```

## Illustrative Explanation

Why good yield prediction does not necessarily mean good EONR prediction?

```{r illustrate-explain, fig.cap="Estimated yield response curves at a specific location", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "802_yield.png"))
```

- Yield level prediction
- Yield response (marginal effect, treatment effect) prediction $\Rightarrow$ More important

## Illustrative Explanation (cont'd)

Estimated vs. True treatment effect

```{r illustrate-treatment-effect, fig.cap="", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "treatment_sim_58.png"))
```

## Field Sizes

```{r comb-boxplot-all, fig.cap="The performances of different models, large field.", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
(comb_boxplot_large + ggtitle("(A) large field")) / (comb_boxplot_medium + ggtitle("(B) medium field")) / (comb_boxplot_small + ggtitle("(C) small field"))
```

- Machine learning models' performances catch up when field sizes decrease
- BRF beats LM, and shrinks gaps to SER, for smaller fields
- Performances of Traditional regression models are more sensitive to field sizes
- Machine learning models are more suitable for small fields

## Efficiency is Money

- At the presence of unknown spatial heterogeneity,
- SER model improve efficiency of LM estimates,
- But it is only treated as a statistical property
- Our simulation evidence: Statistical efficiency means more profitable EONR recommendations

```{r, fig.cap="The performances of different models, large field.", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
(comb_boxplot_large + ggtitle("(A) large field")) / (comb_boxplot_medium + ggtitle("(B) medium field")) / (comb_boxplot_small + ggtitle("(C) small field"))
```

## Conclusions

- Machine learning models perform poor in EONR prediction compared to traditional regression models, despite their better yield prediction.

- What matters more is estimating yield responses, not yield levels.

- SER model can generate the most profitable EONR recommendations for normal-sized fields

- Machine learning models are more useful for small fields.

- More actively use SER model in contrast to LM in yield response estimation.