### 2022-02-26 Writing/appendix.rmd
codes to generate pairwise t-test updated

# Pairwise t-test of yield RMSE, EONR RMSE, and profit between models {-}
```{r count-table, echo=FALSE, cache=FALSE}

prt_table <-
  est_data %>%
  .[order(sim, model), ] %>%
  .[, y := (rmse_cv == min(rmse_cv)), by = .(sim, field_size)] %>%
  .[, eonr := (rmse_eonr == min(rmse_eonr)), by = .(sim, field_size)] %>%
  .[, profit := (profit == max(profit)), by = .(sim, field_size)] %>%
  .[,
    .(
      yield = (mean(y) * 100) %>% round(2),
      EONR = (mean(eonr) * 100) %>% round(2),
      profit = (mean(profit) * 100) %>% round(2)
    ),
    by = .(field_size, model)
  ]

#* create a table
count_table <-
  prt_table %>%
  # === Wide to Long: melt()
  melt(id.vars = c("field_size", "model")) %>%
  # === Long to Wide: dcast()
  dcast(field_size + variable ~ model, value.var = "value")

#* displary the table
knitr::kable(count_table,
  format = "markdown",
  caption = "Percentage (%) of each model being the best-performing among the five models in a single simulation round"
)
```

```{r run-pairwise-test, inlcude = FALSE}
est_data_by_model_1 <-
  est_data %>%
  arrange(sim) %>%
  nest_by(model, field_col) %>%
  mutate(data = list(
    data.table(data)
  )) %>%
  setnames(names(.), paste0(names(.), "_1"))

est_data_by_model_2 <-
  copy(est_data_by_model_1) %>%
  setnames(names(.), gsub("_1", "_2", names(.)))

source(here("GitControlled/Codes/Functions/R/expand_grid_df.R"))

test_data <-
  expand_grid_df(est_data_by_model_1, est_data_by_model_2) %>%
  #* different models
  filter(model_1 != model_2) %>%
  #* same field size
  filter(field_col_1 == field_col_2) %>%
  mutate(yield_pt = list(
    t.test(
      data_1[, rmse_cv],
      data_2[, rmse_cv],
      paired = TRUE
    )
  )) %>%
  mutate(eonr_pt = list(
    t.test(
      data_1[, rmse_eonr],
      data_2[, rmse_eonr],
      paired = TRUE
    )
  )) %>%
  mutate(pi_pt = list(
    t.test(
      data_1[, profit],
      data_2[, profit],
      paired = TRUE
    )
  ))
```

```{r pair-table-yield, tab.cap = "Pairwise t-test of predicted yield RMSE between models"}

test_data %>%
  select(field_col_1, model_1, model_2, yield_pt) %>%
  mutate(
    yield_pt_txt =
      paste0(
        round(yield_pt$estimate, digits = 2), " (",
        round(yield_pt$statistic, digits = 2), ")"
      )
  ) %>%
  data.table() %>%
  dcast(field_col_1 + model_1 ~ model_2, value.var = "yield_pt_txt") %>%
  .[model_1 != "SER", ] %>%
  .[, RF := NULL] %>%
  .[model_1 == "BRF", `:=`(RF_b = NA)] %>%
  .[model_1 == "OLS", `:=`(BRF = NA, RF_b = NA)] %>%
  setnames(c("field_col_1", "model_1"), c("Field Size", "Model\\Model")) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  hline(i = c(4, 8)) %>%
  autofit() %>%
  fix_border_issues() %>%
  align(j = c(1, 2), align = "center")
```

```{r pair-table-eonr, tab.cap = "Pairwise t-test of predicted EONR RMSE between models"}

test_data %>%
  select(field_col_1, model_1, model_2, eonr_pt) %>%
  mutate(
    eonr_pt_txt =
      paste0(
        round(eonr_pt$estimate, digits = 2), " (",
        round(eonr_pt$statistic, digits = 2), ")"
      )
  ) %>%
  data.table() %>%
  dcast(field_col_1 + model_1 ~ model_2, value.var = "eonr_pt_txt") %>%
  .[model_1 != "SER", ] %>%
  .[, RF := NULL] %>%
  .[model_1 == "BRF", `:=`(RF_b = NA)] %>%
  .[model_1 == "OLS", `:=`(BRF = NA, RF_b = NA)] %>%
  setnames(c("field_col_1", "model_1"), c("Field Size", "Model\\Model")) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  hline(i = c(4, 8)) %>%
  autofit() %>%
  fix_border_issues() %>%
  align(j = c(1, 2), align = "center")

# Note: values are the difference in RMSE (row vs. index) (kg/ha), p-values in parenthese

```

```{r pair-table-profit, tab.cap = "Pairwise t-test of profits between models"}

test_data %>%
  select(field_col_1, model_1, model_2, pi_pt) %>%
  mutate(
    pi_pt_txt =
      paste0(
        round(pi_pt$estimate, digits = 2), " (",
        round(pi_pt$statistic, digits = 2), ")"
      )
  ) %>%
  data.table() %>%
  dcast(field_col_1 + model_1 ~ model_2, value.var = "pi_pt_txt") %>%
  .[model_1 != "SER", ] %>%
  .[, RF := NULL] %>%
  .[model_1 == "BRF", `:=`(RF_b = NA)] %>%
  .[model_1 == "OLS", `:=`(BRF = NA, RF_b = NA)] %>%
  setnames(c("field_col_1", "model_1"), c("Field Size", "Model\\Model")) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  hline(i = c(4, 8)) %>%
  autofit() %>%
  fix_border_issues() %>%
  align(j = c(1, 2), align = "center")
```